\textbf{SPECFEM3D}\\ \textbf{Cartesian}\\ \textbf{User Manual}

### 2023-11-17 doc/USER_MANUAL/05_running_the_solver.tex
- added Brune and Smoothed Brune source time functions

\noindent The \texttt{FORCESOLUTION} file should be edited in the
following way:
\begin{itemize}
\item The first line is only the header for the force solution, which can
be used as the identifier for the force source.
\item \texttt{time shift:} For a single force source, set
this  parameter equal to $0.0$ (The solver will not run otherwise!); the time
shift parameter would simply apply an overall time shift to the synthetics,
something that can be done in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item \texttt{hdurorf0:} Set the half duration value (s) for step source-time
function, or the dominant frequency (Hz) for Ricker source-time functions.
In case that the source uses a (pseudo) Dirac delta source-time function (i.e., a Gaussian with a zero half duration) to represent
a force point source, a very short half duration of five time steps is automatically set by default.
For a Ricker source-time function, set the dominant frequency value (f0) in Hz.
See the parameter \texttt{source time function:} below for source-time functions.
\item \texttt{latorUTM:} Set the latitude or UTM $x$ coordinate.
\item \texttt{longorUTM:} Set the longitude or UTM $y$ coordinate.
\item \texttt{depth:} Set the depth of the source (in km).
\item \texttt{source time function:} Set the type of source-time function: 0 = Gaussian function, 1 = Ricker function, 2 = Heaviside (step) function, 3 = monochromatic function, 4 = Gaussian function as defined in \citet{Meschede2011}, 5 = Brune function, and 6 = Smoothed Brune function. For the Brune and Smoothed Brune source time functions \texttt{hdurorf0} is the rise time. The Brune and Smoothed Brune functions are currently implemented only for the viscoelastic simulations.
When {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on in the main parameter file \texttt{DATA/Par\_file},
it will override this source time function type selection and always use a Ricker wavelet.
Note that we use the standard definition of a Ricker, for a dominant frequency $f_0$:
$\mathrm{Ricker}(t) = (1 - 2 a t^2) e^{-a t^2}$, with $a = \pi^2 f_0^2$,
whose Fourier transform is thus:
$\frac{1}{2} \frac{\sqrt{\pi}\omega^2}{a^{3/2}}e^{-\frac{\omega^2}{4 a}}$
This gives the wavelet of Figure~\ref{fig:RickerWavelet}.
\item \texttt{factor force source:} Set the magnitude of the force source (units in Newton N).
\item \texttt{component dir vect source E:} Set the East component of a direction vector for the force
source. Direction vector is not necessarily a unit vector.
\item \texttt{component dir vect source N:} Set the North component of a direction vector for the force
source.
\item \texttt{component dir vect source Z\_up:} Set the vertical component of a direction vector for the force
source. Sign convention follows the positive upward direction (see Appendix A for the orientation
of the reference frame).
\end{itemize}
\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
At least one entry (not necessarily the first) must have a zero \texttt{time
shift}, and all the other entries must have non-negative \texttt{time shift}.
Each subevent can have its own set of parameters \texttt{latitude}, \texttt{longitude}, \texttt{depth},
\texttt{half duration} etc.\newline

>>>>

\noindent The \texttt{FORCESOLUTION} file should be edited in the
following way:
\begin{itemize}
\item The first line is only the header for the force solution, which can
be used as the identifier for the force source.
\item \texttt{time shift:} For a single force source, set
this  parameter equal to $0.0$ (The solver will not run otherwise!); the time
shift parameter would simply apply an overall time shift to the synthetics,
something that can be done in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item \texttt{hdurorf0:} Set the half duration value (s) for step source-time
function, or the dominant frequency (Hz) for Ricker source-time functions.
In case that the source uses a (pseudo) Dirac delta source-time function (i.e., a Gaussian with a zero half duration) to represent
a force point source, a very short half duration of five time steps is automatically set by default.
For a Ricker source-time function, set the dominant frequency value (f0) in Hz.
See the parameter \texttt{source time function:} below for source-time functions.
\item \texttt{latorUTM:} Set the latitude or UTM $x$ coordinate.
\item \texttt{longorUTM:} Set the longitude or UTM $y$ coordinate.
\item \texttt{depth:} Set the depth of the source (in km).
\item \texttt{source time function:} Set the type of source-time function: 0 = Gaussian function, 1 = Ricker function, 2 = Heaviside (step) function, 3 = monochromatic function, 4 = Gaussian function as defined in \citet{Meschede2011}, 5 = Brune function, and 6 = Smoothed Brune function. Please note that we have implemented time derivatives of the Brune and smoothed Brune functions as the moment rate functions. For these source time functions, \texttt{hdurorf0} is the source duration or the rise time. The Brune and Smoothed Brune functions are currently implemented only for the viscoelastic simulations.
When {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on in the main parameter file \texttt{DATA/Par\_file},
it will override this source time function type selection and always use a Ricker wavelet.
Note that we use the standard definition of a Ricker, for a dominant frequency $f_0$:
$\mathrm{Ricker}(t) = (1 - 2 a t^2) e^{-a t^2}$, with $a = \pi^2 f_0^2$,
whose Fourier transform is thus:
$\frac{1}{2} \frac{\sqrt{\pi}\omega^2}{a^{3/2}}e^{-\frac{\omega^2}{4 a}}$
This gives the wavelet of Figure~\ref{fig:RickerWavelet}.
\item \texttt{factor force source:} Set the magnitude of the force source (units in Newton N).
\item \texttt{component dir vect source E:} Set the East component of a direction vector for the force
source. Direction vector is not necessarily a unit vector.
\item \texttt{component dir vect source N:} Set the North component of a direction vector for the force
source.
\item \texttt{component dir vect source Z\_up:} Set the vertical component of a direction vector for the force
source. Sign convention follows the positive upward direction (see Appendix A for the orientation
of the reference frame).
\end{itemize}
\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
At least one entry (not necessarily the first) must have a zero \texttt{time
shift}, and all the other entries must have non-negative \texttt{time shift}.
Each subevent can have its own set of parameters \texttt{latitude}, \texttt{longitude}, \texttt{depth},
\texttt{half duration} etc.\newline
